### 2019-04-16 reports/article.tex
reorganization of notebook and figure names + fix landscape plots

% The evolution of the urban agglomeration
The spatiotemporal evolution of the urban footprint for the three urban agglomerations is displayed in \autoref{figures/landscapes.pdf}.

\begin{figure}[!ht]
  \begin{adjustwidth}{-.4\textwidth}{0cm}
    \centering  
    \includegraphics[width=\linewidth]{figures/landscapes.pdf}
    \vspace{.5em}
    \caption[Evolution of urban patches]{\label{figures/landscapes.pdf}Evolution of urban patches of the three urban agglomerations throughout their respective periods of study. The times t$_0$, t$_1$, t$_2$ and t$_3$ correspond to 1981, 1993, 2004 and 2013 for Bern; 1980, 1990, 2005 and 2014 for Lausanne and 1982, 1994, 2007 and 2016 for Zurich}
  \end{adjustwidth}
\end{figure}

@@

% TODO: Start by summary
% Mostly monotonic trends consistent with coalescence stages; Bern and Lausanne almost identical, Zurich shows similar trend but with urbanization being more complete (more PLAND, MPS and LPI, also more edginess and complexity ED, AWMFD, LSI, more connected ENN and CONTAG and more diverse SHDI)
The computed time series of landscape metrics for the agglomerations of Bern, Lausanne and Zurich are displayed in \autoref{figures/metrics.pdf}.
% TODO: computed at the agglomeration scale. The sensitivity to the extent has been tested in the supplmentary materials.
% TODO: as suggested by the demographic trends
% AREA/SHAPE/EDGE
The proportion of landscape occupied by urban patches, represented by PLAND has increased monotonically for the three agglomerations. Bern and Lausanne show almost indistinguishable trends, starting from a 13$\%$ in the early 1980s and surpass the 16$\%$ in the last snapshot of 2013 and 2014 respectively. Zurich shows a parallel tendence with the percentage of urbanized land increasing from 22$\%$ in 1982 to a 27$\%$ in 2016.
The mean area of urban patches and the percentage of area occupied by the largest urban patch, represented respectively by the MPS and LPI metrics, also show a monotonic increase, parallel among the three agglomerations, but again with higher values for Zurich. % TODO: consistent with coalescence, contradicting diversity, fragmentation and complexity hypothesis.
On the other hand, the density of urban patches, represented by PD, reveals more complex and idiosyncratic trends. % which do not consistently match any of the hypothesis
A generic tendence to decrease first and increase later might be noted, nevertheless Lausanne undergoes a slight decrease during the last period. 
The Bern agglomeration shows the highest values, which is also consistent with the agglomeration having the lowest values in MPS and LPI, denoting a landscape configured by numerous but small urban patches.
Overall, the increases at the latter stages suggest the emergence of new urban patches in a leapfrog manner.
Despite the irregular trends of PD, the edge density (ED) shows more consistent tendencies of increase in Bern and Lausanne and decrease in Zurich. From the perspective of the diffusion and coalescence hypothesis, Bern and Lausanne are seemingly undergoing a diffusion stage, whereas the decrease of Zurich is more characteristic of coalescence. % Decrease in Zurich contradicts the diversity, fragmentation and complexity as well as the three growth modes hypothesis.
The flat evolution of ED during the first period in Zurich suggests a transition from diffusion to coalescence, which is also observed in Lausanne during the last period.
This postulate is also supported by the trend of AWMFD, which in Zurich is almost identical to that of ED, and shows an unimodal pattern for Lausanne, also characteristic of a transition from diffusion to coalescence. The decline of the AWMFD observed during the last period in Bern also suggest that albeit tardier than its counterparts, Bern might also be undergoing a shift towards coalescence.
% The unimodal trend AWMFD is also consistent with the three growth modes hypothesis.
The monotonic decrease of the landscape shape index (LSI) reveals that urban patches are becoming increasingly aggregated, which is also characteristic of coalescence. Similarly, the average euclidean nearest-neighbor distance (ENN) shows an overall monotonic decrease which suggests that urban patches are becoming increasingly close in space. % TODO: decrease of LSI contradicts the second group of hypothesis while ENN is consistent with them.

@@

\begin{figure}[ht]
  \begin{adjustwidth}{-.4\textwidth}{0cm}
    \centering  
    \includegraphics[width=\linewidth]{figures/metrics.pdf}
    \vspace{.5em}
    \caption[Time series of landscape metrics]{\label{figures/metrics.pdf}Time series of landscape metrics. The eight metrics of the two upper rows are computed at the urban class level, that is, aggregating the values computed for each urban patch. The two metrics of the lower row are computed at the landscape level, that is, considering the patches of all the classes present within the landscape (urban and non-urban in this case)}
  \end{adjustwidth}
\end{figure}

@@

The area-radius scaling and the patch size-frequency distribution of the three urban agglomerations at each temporal snapshot are plotted in \autoref{figures/bifractal_radial_dimension.pdf} and \autoref{figures/self_organized_criticality.pdf} respectively.

\begin{figure}[ht]
  \begin{adjustwidth}{-.4\textwidth}{0cm}
    \centering
    \includegraphics[width=\linewidth]{figures/bifractal_radial_dimension.pdf}
    \vspace{.5em}
    \caption[Area-radius scaling]{\label{figures/bifractal_radial_dimension.pdf}Area-radius scaling of the three urban agglomerations at each temporal snapshot. The relationship has been estimated by computing the total area occupied by urban land uses laying within a series of radius values (noted by the cross-shaped markers) from 1000 to 20000m, successively increasing by a step of 1000m. The reference center point has been manually retrieved from the OpenStreetMap\footnotemark} % By querying the coordinates of the Node element with the tags `name={city name}` and `admin_centre:4=yes`
  \end{adjustwidth}
\end{figure}

@@

\begin{figure}[ht]
  \begin{adjustwidth}{-.4\textwidth}{0cm}
    \centering
    \includegraphics[width=\linewidth]{figures/self_organized_criticality.pdf}
    \vspace{.5em}
    \caption[Patch size-frequency distribution]{\label{figures/self_organized_criticality.pdf}Patch size-frequency distribution for the three urban agglomerations  at each temporal snapshot}
  \end{adjustwidth}
\end{figure}

@@

% You can use the \nameref{label} command to cite supporting items in the text.
\subsection*{Code S1}
\label{code-sensitivity}
Exploration of the sensitivity of the time series of landscape metrics to the spatial extent of the urban agglomerations, as Jupyter Notebook (IPYNB).
\url{https://github.com/martibosch/swiss-urbanization/blob/master/notebooks/sensitivity-extent.ipynb}

>>>>

% The evolution of the urban agglomeration
The spatiotemporal evolution of the urban footprint for the three urban agglomerations is displayed in \autoref{figures/landscape_plots.pdf}.

\begin{figure}[!ht]
  \begin{adjustwidth}{-.4\textwidth}{0cm}
    \centering  
    \includegraphics[width=\linewidth]{figures/landscape_plots.pdf}
    \vspace{.5em}
    \caption[Evolution of urban patches]{\label{figures/landscape_plots.pdf}Evolution of urban patches of the three urban agglomerations throughout their respective periods of study. The times t$_0$, t$_1$, t$_2$ and t$_3$ correspond to 1981, 1993, 2004 and 2013 for Bern; 1980, 1990, 2005 and 2014 for Lausanne and 1982, 1994, 2007 and 2016 for Zurich}
  \end{adjustwidth}
\end{figure}

@@

% TODO: Start by summary
% Mostly monotonic trends consistent with coalescence stages; Bern and Lausanne almost identical, Zurich shows similar trend but with urbanization being more complete (more PLAND, MPS and LPI, also more edginess and complexity ED, AWMFD, LSI, more connected ENN and CONTAG and more diverse SHDI)
The computed time series of landscape metrics for the agglomerations of Bern, Lausanne and Zurich are displayed in \autoref{figures/metrics_time_series.pdf}.
% TODO: computed at the agglomeration scale. The sensitivity to the extent has been tested in the supplmentary materials.
% TODO: as suggested by the demographic trends
% AREA/SHAPE/EDGE
The proportion of landscape occupied by urban patches, represented by PLAND has increased monotonically for the three agglomerations. Bern and Lausanne show almost indistinguishable trends, starting from a 13$\%$ in the early 1980s and surpass the 16$\%$ in the last snapshot of 2013 and 2014 respectively. Zurich shows a parallel tendence with the percentage of urbanized land increasing from 22$\%$ in 1982 to a 27$\%$ in 2016.
The mean area of urban patches and the percentage of area occupied by the largest urban patch, represented respectively by the MPS and LPI metrics, also show a monotonic increase, parallel among the three agglomerations, but again with higher values for Zurich. % TODO: consistent with coalescence, contradicting diversity, fragmentation and complexity hypothesis.
On the other hand, the density of urban patches, represented by PD, reveals more complex and idiosyncratic trends. % which do not consistently match any of the hypothesis
A generic tendence to decrease first and increase later might be noted, nevertheless Lausanne undergoes a slight decrease during the last period. 
The Bern agglomeration shows the highest values, which is also consistent with the agglomeration having the lowest values in MPS and LPI, denoting a landscape configured by numerous but small urban patches.
Overall, the increases at the latter stages suggest the emergence of new urban patches in a leapfrog manner.
Despite the irregular trends of PD, the edge density (ED) shows more consistent tendencies of increase in Bern and Lausanne and decrease in Zurich. From the perspective of the diffusion and coalescence hypothesis, Bern and Lausanne are seemingly undergoing a diffusion stage, whereas the decrease of Zurich is more characteristic of coalescence. % Decrease in Zurich contradicts the diversity, fragmentation and complexity as well as the three growth modes hypothesis.
The flat evolution of ED during the first period in Zurich suggests a transition from diffusion to coalescence, which is also observed in Lausanne during the last period.
This postulate is also supported by the trend of AWMFD, which in Zurich is almost identical to that of ED, and shows an unimodal pattern for Lausanne, also characteristic of a transition from diffusion to coalescence. The decline of the AWMFD observed during the last period in Bern also suggest that albeit tardier than its counterparts, Bern might also be undergoing a shift towards coalescence.
% The unimodal trend AWMFD is also consistent with the three growth modes hypothesis.
The monotonic decrease of the landscape shape index (LSI) reveals that urban patches are becoming increasingly aggregated, which is also characteristic of coalescence. Similarly, the average euclidean nearest-neighbor distance (ENN) shows an overall monotonic decrease which suggests that urban patches are becoming increasingly close in space. % TODO: decrease of LSI contradicts the second group of hypothesis while ENN is consistent with them.

@@

\begin{figure}[ht]
  \begin{adjustwidth}{-.4\textwidth}{0cm}
    \centering  
    \includegraphics[width=\linewidth]{figures/metrics_time_series.pdf}
    \vspace{.5em}
    \caption[Time series of landscape metrics]{\label{figures/metrics_time_series.pdf}Time series of landscape metrics. The eight metrics of the two upper rows are computed at the urban class level, that is, aggregating the values computed for each urban patch. The two metrics of the lower row are computed at the landscape level, that is, considering the patches of all the classes present within the landscape (urban and non-urban in this case)}
  \end{adjustwidth}
\end{figure}

@@

The area-radius scaling and the patch size-frequency distribution of the three urban agglomerations at each temporal snapshot are plotted in \autoref{figures/area_radius_scaling.pdf} and \autoref{figures/size_frequency_distribution.pdf} respectively.

\begin{figure}[ht]
  \begin{adjustwidth}{-.4\textwidth}{0cm}
    \centering
    \includegraphics[width=\linewidth]{figures/area_radius_scaling.pdf}
    \vspace{.5em}
    \caption[Area-radius scaling]{\label{figures/area_radius_scaling.pdf}Area-radius scaling of the three urban agglomerations at each temporal snapshot. The relationship has been estimated by computing the total area occupied by urban land uses laying within a series of radius values (noted by the cross-shaped markers) from 1000 to 20000m, successively increasing by a step of 1000m. The reference center point has been manually retrieved from the OpenStreetMap\footnotemark} % By querying the coordinates of the Node element with the tags `name={city name}` and `admin_centre:4=yes`
  \end{adjustwidth}
\end{figure}

@@

\begin{figure}[ht]
  \begin{adjustwidth}{-.4\textwidth}{0cm}
    \centering
    \includegraphics[width=\linewidth]{figures/size_frequency_distribution.pdf}
    \vspace{.5em}
    \caption[Patch size-frequency distribution]{\label{figures/size_frequency_distribution.pdf}Patch size-frequency distribution for the three urban agglomerations  at each temporal snapshot}
  \end{adjustwidth}
\end{figure}

@@

% You can use the \nameref{label} command to cite supporting items in the text.
\subsection*{Code S1}
\label{code-sensitivity}
Exploration of the sensitivity of the time series of landscape metrics to the spatial extent of the urban agglomerations, as Jupyter Notebook (IPYNB).
\url{https://github.com/martibosch/swiss-urbanization/blob/master/notebooks/sensitivity_extent.ipynb}
